Quantizing the Bosonic String on a Loop Quantum Gravity Background

### 2024-06-22 lqg-strings.tex
made changes in section V and added a table with values for the normal-ordering constant

As usual, we define the vacuum state of the string to obey
\begin{equation} \label{eqn:annihilation-vacuum}
	\Hat{a}^{\mu} _{p} \ket{0} = 0 , \ \mathrm{for} \ p \neq 0 .
\end{equation}
For $p = 0$ we have center of mass position and momentum, so the vacuum has extra structure obeying
\begin{subequations}
	\begin{align}
		\Hat{x}^{\mu} \ket{0 ; x} & = x^{\mu} \ket{0 ; x} \label{eqn:x-vacuum-x}                       \\
		\Hat{p}_{\mu} \ket{0 ; x} & = - i \frac{\del}{\del x^{\mu}} \ket{0 ; x} \label{eqn:p-vacuum-x}
	\end{align}
\end{subequations}
in position representation and
\begin{subequations}
	\begin{align}
		\Hat{x}^{\mu} \ket{0 ; p} & = - i \frac{\del}{\del p_{\mu}} \ket{0 ; p} \label{eqn:x-vacuum-p} \\
		\Hat{p}_{\mu} \ket{0 ; p} & = p_{\mu} \ket{0 ; p} \label{eqn:p-vacuum-p}
	\end{align}
\end{subequations}
in momentum representation.

@@

When translating the constraints as quantum operators, we require that they have vanishing matrix elements when sandwiched by physical states $\psi$ and $\phi$:
\begin{subequations}
	\begin{equation} \label{eqn:L-constraint-quantum-1}
		\bra{\phi} \Hat{L}_{p , k} \ket{\psi} = 0 \, \,
	\end{equation}
	\begin{equation} \label{eqn:L-constraint-quantum-2}
		\bra{\phi} \Hat{\widetilde{L}}_{p , k} \ket{\psi} = 0 .
	\end{equation}
\end{subequations}
$\Hat{L}_{p , k}$ has no ambiguities since it's composed of only annihilation operators. As for $\widetilde{L}_{p , k}$, we have an ambiguity for $p = k \neq 0$, so we pick normal ordering with the annihilation operator to the right
\begin{equation} \label{eqn:L-normal-ordering}
	\Hat{\widetilde{L}}_{p , p} = (\Hat{\alpha}^{\mu} _{p})^{\dagger} \Hat{\alpha}^{\nu} _{p} \eta _{\mu \nu} .
\end{equation}
This ambiguity manifests in the imposition of the constraint as
\begin{equation} \label{eqn:shifted-L-constraint}
	\bra{\phi} \left( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 \Hat{\widetilde{L}}_{n , n} \right) - a \right) \ket{\psi} = 0 ,
\end{equation}
for some constant $a$. Classically, we had eq (\ref{eqn:string-rest-mass-classical-1}) for the string rest mass, from which we can now see that the string mass spectrum will get shifted by
\begin{multline} \label{eqn:shifted-mass-spectrum}
	\Hat{M}^2 = \frac{8}{\alpha '} \Bigg( \frac{2 \mu ^2}{\alpha '} \Hat{x}^{\mu} \Hat{x}_{\mu} - \\
	- \left( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 E_{n} (\Hat{a}^{\mu} _{n})^{\dagger} \Hat{a}^{\nu} _{n} \eta _{\mu \nu} \right) - a \right) \Bigg) .
\end{multline}
This constant can be calculated by imposing normal-ordering on the summed constraint since $n \in (- d , d)$ is finite, and the result is
\begin{multline}
	a = - D \sum _{n \neq 0} \left( 1 + \frac{n}{E_{n}} \right) ^2 E_{n} = \\
	= - D \sum _{n > 0} \left( 1 + \frac{n^2}{E_{n} ^2} \right) E_{n} .
\end{multline}

The commutation relations between the constraints are inherited by the Poisson brackets
\begin{subequations}
	\begin{align}
		\left[ \Hat{L}_{n , m} , \Hat{L}_{k , p} \right]                         & = 0 \label{eqn:L-commutator-0}                                                                                                   \\
		\left[ \Hat{L}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right]             & = \Hat{L}_{m , k} \delta _{|n| , |p|} + \Hat{L}_{n , k} \delta _{|m| , |p|} \label{eqn:L-commutator-1}                           \\
		\left[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] & = \Hat{\widetilde{L}}_{k , m} \delta _{|n| , |p|} - \Hat{\widetilde{L}}_{n , p} \delta _{|m| , |k|} \label{eqn:L-commutator-2} .
	\end{align}
\end{subequations}
The commutator (\ref{eqn:L-commutator-1}) ends up with only annihilation operators in the R.H.S, so it has no ordering ambiguities, meanwhile the commutator (\ref{eqn:L-commutator-2}) have ordering ambiguities for $|m| = |k|$ and/or $|n| = |p|$, so we add the anomalous terms
\begin{multline} \label{eqn:L-commutator-anomalous-initial}
	\left[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] = \Hat{\widetilde{L}}_{k , m} \delta _{|n| , |p|} - \Hat{\widetilde{L}}_{n , p} \delta _{|m| , |k|} + \\
	+ C_{n} \delta _{|n| , |p|} + D_{k} \delta _{|m| , |k|} .
\end{multline}
Clearly, $C_{0} = D_{0} = 0$, since when all indices are $0$ we have only the position and momentum operators, and thus the commutator should vanish identically. Also, because of the symmetry in the indices we have that $C_{- n} = C_{n}$ and $D_{- k} = D_{k}$. Using the Jacobi identity we get the difference equation
\begin{equation} \label{eqn:F-difference-eqn}
	F_{n + 1} + F_{n} + F_{1} = 0
\end{equation}
for the sum of both sequences $F_{n} = C_{n} + D_{n}$. This difference equation has solution
\begin{equation} \label{eqn:F-sequence}
	F_{n} = \frac{F_{1}}{2} \left( \cos \left( \frac{n \pi}{2} \right) - 1 \right) ,
\end{equation}
and since both $C_{n}$ and $D_{n}$ need to satisfy the initial value of $0$, each of then is half of $F_{n}$,
\begin{equation} \label{eqn:C-D-sequences}
	C_{n} = D_{n} = \frac{F_{n}}{2} = \frac{F_{1}}{4} \left( \cos \left( \frac{n \pi}{2} \right) - 1 \right) .
\end{equation}
Next, by calculating the vacuum expectation value (VEV) of the commutator \eqref{eqn:L-commutator-2} with $|m| \neq |k|$ and $n = p = 1$, we find that $F_{1} = 0$, thus both sequences $C_{n} = D_{n} = 0$ vanish and we conclude that the quantum version of our algebra \emph{has no anomalies}. This is a relevant result, since it could mean our model doesn't need a specific number of dimensions to work, which will be explored again later in section \ref{sec:discussion}.

>>>>

As usual, we define the vacuum state of the string to obey
\begin{equation} \label{eqn:annihilation-vacuum}
	\Hat{a}^{\mu} _{p} \ket{0} = 0 , \ \mathrm{for} \ p \neq 0 .
\end{equation}
For $p = 0$ we have center of mass position and momentum, so the vacuum has extra structure obeying
\begin{subequations}
	\begin{align}
		\Hat{x}^{\mu} \ket{0 ; x} & = x^{\mu} \ket{0 ; x} \label{eqn:x-vacuum-x}                       \\
		\Hat{p}_{\mu} \ket{0 ; x} & = - i \frac{\del}{\del x^{\mu}} \ket{0 ; x} \label{eqn:p-vacuum-x}
	\end{align}
\end{subequations}
in position representation $\ket{0 ; x} \equiv \ket{0} \otimes \psi (x)$ and
\begin{subequations}
	\begin{align}
		\Hat{x}^{\mu} \ket{0 ; p} & = - i \frac{\del}{\del p_{\mu}} \ket{0 ; p} \label{eqn:x-vacuum-p} \\
		\Hat{p}_{\mu} \ket{0 ; p} & = p_{\mu} \ket{0 ; p} \label{eqn:p-vacuum-p}
	\end{align}
\end{subequations}
in momentum representation $\ket{0 ; p} \equiv \ket{0} \otimes \psi (p)$.

@@

When translating the constraints as quantum operators, we impose them strongly, requiring that they have vanishing matrix elements when sandwiched by physical states $\psi$ and $\phi$ since said constraints are not self-adjoint:
\begin{subequations}
	\begin{equation} \label{eqn:L-constraint-quantum-1}
		\bra{\phi} \Hat{L}_{p , k} \ket{\psi} = 0 \, \,
	\end{equation}
	\begin{equation} \label{eqn:L-constraint-quantum-2}
		\bra{\phi} \Hat{\widetilde{L}}_{p , k} \ket{\psi} = 0 .
	\end{equation}
\end{subequations}
$\Hat{L}_{p , k}$ has no ambiguities since it's composed of only annihilation operators. As for $\widetilde{L}_{p , k}$, we have an ambiguity for $p = k \neq 0$, so we pick normal ordering with the annihilation operator to the right
\begin{equation} \label{eqn:L-normal-ordering}
	\Hat{\widetilde{L}}_{p , p} = (\Hat{\alpha}^{\mu} _{p})^{\dagger} \Hat{\alpha}^{\nu} _{p} \eta _{\mu \nu} .
\end{equation}
This ambiguity manifests in the imposition of the constraint as
\begin{equation} \label{eqn:shifted-L-constraint}
	\bra{\phi} \left( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 \Hat{\widetilde{L}}_{n , n} \right) - a \right) \ket{\psi} = 0 ,
\end{equation}
for some constant $a$. Classically, we had eq (\ref{eqn:string-rest-mass-classical-1}) for the string rest mass, from which we can now see that the string mass spectrum will get shifted by this constant
\begin{multline} \label{eqn:shifted-mass-spectrum}
	\Hat{M}^2 = \frac{8}{\alpha '} \Bigg( \frac{2 \mu ^2}{\alpha '} \Hat{x}^{\mu} \Hat{x}_{\mu} - \\
	- \left( \sum _{n \neq 0} \left( \left( 1 + \frac{n}{E_{n}} \right) ^2 E_{n} (\Hat{a}^{\mu} _{n})^{\dagger} \Hat{a}^{\nu} _{n} \eta _{\mu \nu} \right) - a \right) \Bigg) .
\end{multline}
This constant can be calculated by imposing normal-ordering on the summed constraint since $n \in (- d , d)$ is finite, and the result is
\begin{multline}
	a = - D \sum _{n \neq 0} \left( 1 + \frac{n}{E_{n}} \right) ^2 E_{n} = \\
	= - D \sum _{n > 0} \left( 1 + \frac{n^2}{E_{n} ^2} \right) E_{n} ,
\end{multline}
which is easily computable given a value for $\mu$. Below we provide a table with the value of $a$ for different spacetime dimensionalities $D$ and mass parameters $\mu$.
\begin{table}[h] \label{tab:normal-ordering-constant}
	\begin{tabular}{|l|l|l|l|l|l|}
		\hline
		        & $\mu = 1$ & $\mu = 2$ & $\mu = 3$ & $\mu = 4$ & $\mu = 5$ \\ \hline
		$D = 2$ & $-4.24$   & $-5.37$   & $-6.96$   & $-8.73$   & $-10.59$  \\ \hline
		$D = 3$ & $-18.44$  & $-20.78$  & $-24.58$  & $-29.20$  & $-34.27$  \\ \hline
		$D = 4$ & $-48.62$  & $-52.11$  & $-58.23$  & $-66.13$  & $-75.19$  \\ \hline
		$D = 5$ & $-100.79$ & $-105.39$ & $-113.79$ & $-125.09$ & $-138.49$ \\ \hline
		$D = 6$ & $-180.96$ & $-186.63$ & $-197.25$ & $-211.95$ & $-229.84$ \\ \hline
	\end{tabular}
	\caption{values of $a$ for different spacetime dimensionalities and mass parameters. Visibly, the dependence in $D$ is much stronger than the dependence in $\mu$.}
\end{table}

The commutation relations between the constraints are inherited by the Poisson brackets
\begin{subequations}
	\begin{align}
		\left[ \Hat{L}_{n , m} , \Hat{L}_{k , p} \right]                         & = 0 \label{eqn:L-commutator-0}                                                                                                   \\
		\left[ \Hat{L}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right]             & = \Hat{L}_{m , k} \delta _{|n| , |p|} + \Hat{L}_{n , k} \delta _{|m| , |p|} \label{eqn:L-commutator-1}                           \\
		\left[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] & = \Hat{\widetilde{L}}_{k , m} \delta _{|n| , |p|} - \Hat{\widetilde{L}}_{n , p} \delta _{|m| , |k|} \label{eqn:L-commutator-2} .
	\end{align}
\end{subequations}
The commutator (\ref{eqn:L-commutator-1}) ends up with only annihilation operators in the R.H.S, so it has no ordering ambiguities, meanwhile the commutator (\ref{eqn:L-commutator-2}) have ordering ambiguities for $|m| = |k|$ and/or $|n| = |p|$, so we add the anomalous terms
\begin{multline} \label{eqn:L-commutator-anomalous-initial}
	\left[ \Hat{\widetilde{L}}_{n , m} , \Hat{\widetilde{L}}_{k , p} \right] = \Hat{\widetilde{L}}_{k , m} \delta _{|n| , |p|} - \Hat{\widetilde{L}}_{n , p} \delta _{|m| , |k|} + \\
	+ C_{n} \delta _{|n| , |p|} + D_{k} \delta _{|m| , |k|} .
\end{multline}
Clearly, $C_{0} = D_{0} = 0$, since when all indices are $0$ we have only the position and momentum operators, and thus the commutator should vanish identically. Also, because of the symmetry in the indices we have that $C_{- n} = C_{n}$ and $D_{- k} = D_{k}$. Using the Jacobi identity we get the difference equation
\begin{equation} \label{eqn:F-difference-eqn}
	F_{n + 1} + F_{n} + F_{1} = 0
\end{equation}
for the sum of both sequences $F_{n} = C_{n} + D_{n}$. This difference equation has solution
\begin{equation} \label{eqn:F-sequence}
	F_{n} = \frac{F_{1}}{2} \left( \cos \left( \frac{n \pi}{2} \right) - 1 \right) ,
\end{equation}
and since both $C_{n}$ and $D_{n}$ need to satisfy the initial value of $0$, each of then is half of $F_{n}$,
\begin{equation} \label{eqn:C-D-sequences}
	C_{n} = D_{n} = \frac{F_{n}}{2} = \frac{F_{1}}{4} \left( \cos \left( \frac{n \pi}{2} \right) - 1 \right) .
\end{equation}
Next, by calculating the vacuum expectation value (VEV) of the commutator \eqref{eqn:L-commutator-2} with $|m| \neq |k|$ and $n = p = 1$, we find that $F_{1} = 0$, thus both sequences $C_{n} = D_{n} = 0$ vanish and we conclude that the quantum version of our algebra \emph{has no anomalies}. This is a relevant result, since it could mean our model doesn't need a specific number of dimensions to work, which will be explored again later in section \ref{sec:emergent-dimensions}.
